Graphics

### 2019-02-21 contest/graphics/graphics.Rmd
add 2/21 10:30 am version of graphics

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = TRUE)
library(deuce)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(kableExtra)
library(pander)
library(gridExtra)
library(grid)

@@

```{r data-cleaning, include = FALSE}
data("gs")
data("gs_players")
data("gs_partial_pbp")

@@

@@

@@

```{r, fig.height=4, fig.width=12}
france_wins <- filter(gs_players, ioc == "FRA", round < "QF") %>% 
  group_by(tournament, round) %>% 
  summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()))

@@

p3 <- ggplot(aus_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line() + 
  # geom_point() +
  labs(title = "Australian players",
       x = "", y = "Win proportion") + 
  scale_color_manual(" ", values=tournament_colors) + my_theme

grid.arrange(p1,p2, p3, nrow=1, widths = c(.29, .29, .42),
             top = textGrob("Home court advantage debunked: win proportion by country",
                            gp=gpar(fontsize=20)),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           # x = 0,
                           # y = 0.5,
                           just = "left",
                           gp=gpar(fontsize=20, fontfamily="serif")))

@@

@@

@@

```{r, fig.height=3, fig.width=4}
ggplot(gs_players) + geom_point(aes(log(ht+1), log(ace+1), color = factor(did_win)))
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = TRUE)
library(deuce)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(kableExtra)
library(pander)
library(gridExtra)
library(grid)
library(ggrepel)

@@

```{r data-cleaning, include = FALSE}
data("gs")
data("gs_players")
data("gs_partial")

@@

# with yellow for us
tournament_colors <- c("#0297DB", "#b06835", "#ffe500", "#54008b")

@@

```{r, fig.height=5, fig.width=12}
tournament_colors <- c("#0297DB", "#b06835", "#ffe500", "#54008b")

@@

tournament_colors <- c("#0297DB", "#b06835", "#ffd500", "#54008b")

france_wins <- filter(gs_players, ioc == "FRA", round < "QF") %>% 
  group_by(tournament, round) %>% 
  summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()))

@@

p3 <- ggplot(aus_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line() + 
  # geom_point() +
  labs(title = "Australian players",
       x = "", y = "Win proportion") + 
  scale_color_manual(guide=FALSE, values=tournament_colors) + my_theme

p4 <- ggplot(gbr_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  geom_line() + 
  # geom_point() +
  labs(title = "British players",
       x = "", y = "Win proportion") + 
  scale_color_manual(" ", values=tournament_colors) + my_theme

grid.arrange(p3, p1,p2,  p4, nrow=1, widths = c(.22, .22, .22, .34),
             top = textGrob("Home court advantage debunked: win proportion by country",
                            gp=gpar(fontsize=20)),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           # x = 0,
                           # y = 0.5,
                           just = "left",
                           gp=gpar(fontsize=20, fontfamily="serif")))

@@

tournament_colors <- c("#0297DB", "#b06835", "#ffd500", "#54008b")

@@

```{r, fig.width=10, fig.height=10}

p_atp <- gs_players %>% filter(league == "ATP") %>%
ggplot( aes(log(ht+1), log(ace+1))) + 
  geom_jitter(aes(color = tournament)) + 
  scale_color_manual(guide=FALSE, values = tournament_colors) + 
  geom_smooth(aes(color = tournament), method = "lm", se=FALSE) + 
  geom_label_repel(data=subset(gs_players %>% filter(league == "ATP"), (log(ht+1) > 5.25 & log(ace+1) > 3.9) |
                                (log(ht+1) > 5.28 & log(ace+1) < 1) | 
                                (log(ht+1) < 5.2 & log(ace+1) > 3)),
            aes(log(ht+1), log(ace+1), label=name),
            alpha=.6, seed=1234, show.legend=FALSE) + 
  geom_label_repel(data=subset(gs_players, (log(ht+1) > 5.25 & log(ace+1) > 3.9) |
                                (log(ht+1) > 5.28 & log(ace+1) < 1) | 
                                (log(ht+1) < 5.2 & log(ace+1) > 3)),
            aes(log(ht+1), log(ace+1), label = name),
            seed=1234, fill=NA, show.legend=FALSE) +
  labs(x = "log(height + 1)", title = "ATP")+
  my_theme

p_wta <- gs_players %>% filter(league == "WTA") %>%
ggplot(aes(log(ht+1), log(ace+1))) + 
  geom_jitter(aes(color = tournament)) + 
  scale_color_manual("Tournament", values = tournament_colors) + 
  geom_smooth(aes(color = tournament), method = "lm", se=FALSE) + 
  geom_label_repel(data=subset(filter(gs_players, league == "WTA"),
                               (log(ht+1) > 5.22 & log(ace+1) > 2.6) |
                                (log(ht+1) > 5.225 & log(ace+1) < .5) | 
                                (log(ht+1) < 5.13 & log(ace+1) > 2.1) | 
                                (log(ht+1) < 5.09 & log(ace+1) > 1.5) |
                                 (log(ace + 1) > 2.82)),
            aes(log(ht+1), log(ace+1), label=name),
            alpha=.6, seed=1234, show.legend=FALSE) + 
  geom_label_repel(data=subset(filter(gs_players, league == "WTA"),
                               (log(ht+1) > 5.22 & log(ace+1) > 2.6) |
                                (log(ht+1) > 5.225 & log(ace+1) < .5) | 
                                (log(ht+1) < 5.13 & log(ace+1) > 2.1) | 
                                (log(ht+1) < 5.09 & log(ace+1) > 1.5) |
                                 (log(ace + 1) > 2.82)),
            aes(log(ht+1), log(ace+1), label=name),
            seed=1234, fill=NA, show.legend=FALSE) +
  labs(x = "log(height + 1)", title = "WTA")+
  my_theme + theme(legend.position="bottom")

grid.arrange(p_atp, p_wta, nrow=2, heights = c(.46, .54),
             top = textGrob("Taller players are better servers: confirmed",
                            gp=gpar(fontsize=20, fontfamily="serif")))

# dont even know

We find that the distribution of errors and aces differ by tournament. Specifically, GAMES? in Wimbledon follow different ace and error patterns than the other tournaments.

```{r, fig.height=3, fig.width=4}
ggplot(gs_partial, aes(w_n_ue + w_n_ue, w_n_aces + w_n_aces,
                           label = tournament, color=tournament)) + 
  geom_jitter() +
  labs( title = "Distribution of errors and aces differ by tournmanet",
        x = "Total unforced errors", y = "Total aces") +
  my_theme + scale_color_manual("Tournament", values = tournament_colors)

# 5 matches to 3

```{r}
gs_men <- filter(gs, league == "ATP") %>% 
  rowwise() %>%
  mutate(winner_wins_after_3 = sum(W1 > L1,
                                 W2 > L2,
                                 W3 > L3),
         winner_after_3 = if((winner_wins_after_3 >= 2) & 
                             !is.na(winner_wins_after_3)) winner_name else
                               if(is.na(winner_wins_after_3)) NA else loser_name,
         winner_same = ifelse((winner_after_3 == winner_name) & !is.na(winner_after_3),
                              "Yes", ifelse(is.na(winner_after_3), NA, "No")))

table(gs_men$winner_wins_after_3, useNA = "always")
table(is.na(gs_men$winner_after_3))
table(gs_men$winner_same)

prop.table(table(gs_men$winner_same))

ggplot(gs_men) + geom_density(aes(winner_rank - loser_rank, color = factor(winner_same)))

We are interested in how the tournaments differ in terms of number of winners, number of unforced errors, and number of aces for both players within a MATCH???. We

visualize the distance between tournaments

```{r, fig.height=4, fig.width=10, eval=FALSE}
# n_winners_p1, n_winners_p2,
gs_num <- ungroup(filter(gs_partial_pbp, Tour == "atp")) %>%
  select( n_ue_p1, n_ue_p2,
         n_aces_p1, n_aces_p2) %>%
  na.omit()

mds <- cmdscale(dist(scale(gs_num)), k = 2) %>%
  as.data.frame()

# mds <- cmdscale(dist(gs_num), k = 2) %>%
#   as.data.frame() 
mds$tournament <- ungroup(filter(gs_partial_pbp, Tour == "atp")) %>%
  select(n_winners_p1, n_winners_p2, n_ue_p1, n_ue_p2, 
         n_aces_p1, n_aces_p2, tournament) %>% 
  na.omit() %>% pull( tournament)

p_mds <- ggplot(mds, aes(V1, V2, label = tournament, color=tournament)) + 
  geom_jitter() +
  labs( title = "Original") +
  my_theme + scale_color_manual(guide=FALSE, values = tournament_colors)

ggplot(gs_partial_pbp, aes(n_ue_p1 + n_ue_p2, n_aces_p1+ n_aces_p2,
                           label = tournament, color=tournament)) + 
  geom_jitter() +
  labs( title = "Original") +
  my_theme + scale_color_manual( values = tournament_colors)

p_mds_zoom <- ggplot(mds, aes(V1, V2, label = tournament, color=tournament)) + 
  geom_jitter() + xlim(c(-1, 2)) + ylim(c(-2, 1.5)) +
  labs( title = "Zoomed") +
  my_theme + scale_color_manual("Tournament", values = tournament_colors)

grid.arrange(p_mds, p_mds_zoom, nrow=1, widths = c(.42, .58),
             top = textGrob("MDS plot with: aces, winners, and unforced errors",
                            gp=gpar(fontsize=20, fontfamily="serif")))

@@
